The landscape of chloroplast genome assembly tools

### 2019-06-05 manuscript/bmc_article.tex
Moving everything into a subfolder

\usepackage{biocon}
\newplant{At}{genus=Arabidopsis,epithet=thaliana}

@@

% List of docker images
\newcommand{\docker}[2]{\href{https://cloud.docker.com/u/chloroextractorteam/repository/docker/chloroextractorteam/#1}{chloroextracteam/#1:#2}}
\newcommand{\dockerce}{\docker{benchmark\_chloroextractor}{v2.0.0}}
\newcommand{\dockercesha}{\texttt{8f49e03424b37e699c5fea9391f79dbb9f2d0dc550ce86c4c700f39deec2dacd}}
\newcommand{\dockeroa}{\docker{benchmark\_org-asm}{v2.0.0}}
\newcommand{\dockeroasha}{\texttt{ff83677c97b7c4e346191b9e04a5162eeea2a8dfd9e60ff84067d33747868f60}}
\newcommand{\dockerfp}{\docker{benchmark\_fastplast}{v2.0.0}}
\newcommand{\dockerfpsha}{\texttt{a3d06a610f8340ba49c3ff3b27342534e2a5348b17caa4fd3c3f3d327243a272}}
\newcommand{\dockerioga}{\docker{benchmark\_ioga}{v2.0.0}}
\newcommand{\dockeriogasha}{\texttt{4698a21c343c60290bbf16811165a654ff01e8fddeb41cd75d0771b7f45968c0}}
\newcommand{\dockernp}{\docker{benchmark\_novoplasty}{v2.0.0}}
\newcommand{\dockernpsha}{\texttt{106387bad4e8e5c53fb9d4c3bdd60cdddf05a47b52a38697369eeabb947c547c}}
\newcommand{\dockergo}{\docker{benchmark\_getorganelle}{v2.0.0}}
\newcommand{\dockergosha}{\texttt{2ed3a464a82025a196ea56c649d0b0a3472cef76994bb065d56054d93298d956}}
\newcommand{\dockercassp}{\docker{benchmark\_chloroplast\_assembly\_protocol}{v2.0.0}}
\newcommand{\dockercasspsha}{\texttt{d10d2317103d91fadec4cb1d5466ac109f23176ad74b6f44f0179175f050155d}}
\endlocaldefs

@@

\author[
    addressref={aff1,aff3},
    noteref={n1},
    email={jan.freudenthal@uni-wuerzburg.de}
    ]{\inits{JAF}\fnm{Jan A} \snm{Freudenthal}}
\author[
    addressref={aff1,aff2},
    noteref={n1},
    email={simon-pfaff@gmx.de}
    ]{\inits{SP}\fnm{Simon} \snm{Pfaff}}
\author[
   addressref={aff1,aff3},
   email={niklas.terhoeven@uni-wuerzubrg.de}
]{\inits{NT}\fnm{Niklas} \snm{Terhoeven}}
\author[
   addressref={aff1},
   email={arthur.korte@uni-wuerzubrg.de}
]{\inits{AK}\fnm{Arthur} \snm{Korte}}
\author[
   addressref={aff1,aff3,aff2 },
   noteref={n2},
   email={markus.ankenbrand@uni-wuerzubrg.de}
]{\inits{MJA}\fnm{Markus J} \snm{Ankenbrand}}
\author[
   addressref={aff2,aff4},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   noteref={n2},                        % id's of article notes, if any
   email={frank.foerster@uni-wuerzburg.de}   % email address
]{\inits{FF}\fnm{Frank} \snm{Förster}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstractbox}
\begin{abstract} % abstract
%One or two sentences providing a basic introduction to the field, comprehensible to a scientist in any discipline.
Chloroplasts are photosynthetic organelles in plant cells and contain their own genomic information.
That genome can be utilized in different scientific fields like phylogenetics or biotechnology.
%Two to three sentences of more detailed background, comprehensible to scientists in related disciplines.
Thus, different assemblers have been developed specialized in chloroplast assemblies.
Those assemblers often use the output of whole genome sequencing experiments as input.
Such sequencing data usually contain the complete chloroplast genome information, even if the sequencing aims for the core genome.
%One sentence clearly stating the general problem being addressed by this particular study.
Different assembly tools have never been systematically compared. %to each other concerning success rate or computational requirements.
%One sentence summarising the main result (with the words “here we show” or their equivalent).
Here we present a benchmark of seven chloroplast assembly tools, capable to succeed in more than \SI{60}{\percent} of real data sets.
%Two or three sentences explaining what the main result reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.
Our results show significant differences between the tested assemblers in terms of generating whole chloroplast genome sequences and computational requirements.
Moreover, we suggest further development to improve user experience and success rate.
%One or two sentences to put the results into a more general context. 
In terms of reproducibility, we created docker images for each tested tool, which are available for the scientific community.
Following the presented guidelines, users are able to analyse and screen data sets for chloroplast genomes using only standard computer infrastructure.
%Two or three sentences to provide a broader perspective, readily comprehensible to a scientist in any discipline, may be included in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced by their inclusion. 
Thus large scale screening for chloroplasts as hidden treasures within genomic sequencing data is feasible.

@@

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Introduction}
\subsection*{General introduction and motivation}
Chloroplasts are essential organelles present in plant cells and the cells of some protists.
Chloroplasts enable the conversion of light energy into chemical energy via photosynthesis.
They harbor their own ribosomes and a circular DNA genome with a size between \SIrange{120}{160}{\kilo\basepair} \cite{palmer_1985}\todo{diese range umfasst die meisten aber nicht alle Chloroplasten}.
Because of this small size, the chloroplast genome has been an early target for sequencing.
The first chloroplast genome sequences were obtained as early as 1986 \cite{ohyama_chloroplast_1986,shinozaki_complete_1986}.
These early efforts elucidated the general genome organization and structure of the chloroplast DNA. 
Chloroplast genome content and structure are reviewed for example in \cite{wicke_evolution_2011,green_chloroplast_2011}.
Chloroplast genomes are widely used for evolutionary analyses \cite{martin_plastid_2010,xiao-ming_inferring_2017}, barcoding \cite{kress_use_2005,hollingsworth_dna_2009,de_vere_dna_2015}, and meta-barcoding \cite{bell_review_2016,deiner_environmental_2017}.
Interesting aspects of chloroplast genomes are their small size (\SIrange{120}{160}{\kilo\basepair},\cite{palmer_1985}), caused through endosymbiotic gene transfer \cite{martin_evolutionary_2002,timmis_endosymbiotic_2004} and the low number of \numrange{100}{120} genes that are still encoded on the chloroplast genome \cite{wicke_evolution_2011}.
Despite the overall high conservation of the genome sequence, there are striking differences in the gene content between different groups (e.g. the loss of the whole \genename{ndh} gene family in Droseraceae \cite{nevill_plastome-wide_2019}). Even more extreme evolutionary cases, where chloroplasts show a very low GC content and a modified genetic code are described \cite{su_novel_2019}.

@@

Structurally, two inverted repeats (IR\textsubscript{A} and IR\textsubscript{B}) of \SIrange{10}{76}{\kilo\basepair} divide the chloroplast genome into a large (LSC) and a small single copy (SSC) region \cite{palmer_1985}.
Those large inverted repeats complicate automated resolution with short read technologies\cite{Wang2018}.
Moreover, the existence of different chloroplasts within a single individual, and thus multiple different chloroplast genomes, have been described for different plants \cite{corriveau_1988,Chat2002,Scar2016}.
Although the origin and evolutionary importance of this phenomena, - called heteroplasmy - are only poorly understood, the assembly of whole chloroplast genomes might be hindered.

Databases containing short read data for species where no reference chloroplast sequence is publicly available, exist. \todo{cite? SRA?}
The availability of whole chloroplast genomes would enable large scale comparative studies \cite{tonti-filippini_what_2017}.
Additionally, reconstructed full chloroplast genomes have been used as super-barcodes \cite{coissac_barcodes_2016}, for biotechnology applications and genetic engineering \cite{daniell_chloroplast_2016}.

%\subsection*{Chloroplast genome architecture}\ff{Chloroplast genome architecture section}
%
%Maybe separate section describing a prototypical chloroplast: LSC, IRa, SSC, IRb. Gene number and content. 
%Size ranges. 
%Evolutionary history incl. endosymbiontic gene transfer. 
%Reduced genomes in parasitic plants. Loss of \genename{ndh} gene family (maybe too specific).
%Preliminary SNP analysis in \plant{At} to estimate intraspecific variability? Is there anything known about intra-organismic variability (do some plants have multiple populations of chloroplasts? \ak{Yes, they do , Heteroplasmy} probably off topic).\ak{nope, I think this is important}

@@

Different strategies have been developed to assemble chloroplast genomes \cite{twyford_strategies_2017}.
In general, obtaining a chloroplast genome from WGS data requires two steps. First, the chloroplast reads have to be extracted from the mixed sequencing data. 
The second step is the assembly and resolution of the special circular structure including the inverted repeats.
The extraction of the reads can be achieved by mapping the reads to a reference chloroplast. \cite{Vinga2012}.
A different approach that does not perform alignments, relies on the higher coverage of chloroplast data in the whole genome sequencing data set\cite{Chan2013}.
Here, a $k$-mer analysis can be used to extract the most frequent reads. An example for this is implemented in \ce{} \cite{j_ankenbrand_chloroextractor:_2018}. \todo{ce ist hier der einzig explizit genannte Assembler, das wirkt wie bias :) entfernen?}
A third method combines both approaches by useing a reference chloroplast as seed and simultaneously assembling the reads based on $k$-mers \cite{dierckxsens_novoplasty:_2017}.

%\todo{Niklas: wir nennen hier immer nur \ce{} als beispiel - wir sollten auch andere nennen}
%\todo{example}
%\ak{e.g. HomBlocks: A multiple-alignment construction pipeline for organelle phylogenomics based on locally collinear block searching, aber ist nicht inkluded, also können wir zitieren, müssen aber nicht}

@@

\section*{Methods}
\subsection*{Data availability}
Source code for all methods used is available at \url{https://github.com/chloroExtractorTeam/benchmark} and archived in zenodo under \zenodorepo{}.
All docker images are published on \url{https://cloud.docker.com/u/chloroextractorteam/} and are named with a leading \texttt{benchmark\_} (\cref{tab:dockerimages}).

@@

\subsection*{Our Setup}
% done by Frank \todo{ausformulieren}
We want to use a minimum of different parameter settings for all assembly programs to enable a fair comparison. 
Therefore, we decided to specify that all programs have to work based on two input files, representing a data set's forward (\texttt{forward.fq}) and reverse  (\texttt{reverse.fq}) sequence file in FASTQ format.
Depending on the assembler, output files with different names and locations are generated.
Those different files are copied and renamed to ensure that each assembly approach produces the same output file (\texttt{output.fa}). 
Additionally, we set an environment variable for all programs to control the number of allowed threads.
All three requirements (defined input file names, defined output file name, thread number control via environment variable) are ensured by a simple wrapper script (\texttt{wrapper.sh}).
Finally, for a maximum of reproducibility all programs have been bundled into individual docker images based on a central base image which provides all the required software. Those docker images were used for the recording of the consumption of computational resources.
Furthermore, all our docker images have been converted into singularity containers for the quantitative measurement on simulated and real data sets.
Singularity container were built from docker images for usage on a HPC-environment using Singularity v.2.5.2 \cite{kurtzer2017singularity}
All singularity containers were run on %the Julia-HPC 
\ak{I would note refer to Julia in the ms, maybe we could add it as a link}
Intel® Xeon® Gold \num{6140} Processors using a Slurm workload manager version 17.11.8 \cite{Jette02slurm}. Assemblies were run on \num{4} threads using \SI{10}{\gibi\byte} RAM with a time limit of \SI{48}{\hour}. 
\subsection*{Data}
\subsubsection*{Simulated}
To avoid suffering from sequencing errors and biological variances, we simulated perfect reads based on the \plant{At} (TAIR10) chloroplast assembly \cite{tair10}.
We used a sliding window approach with \texttt{seqkit} \cite{seqkit}. The exact commands are documented in \href{https://github.com/chloroExtractorTeam/benchmark/blob/master/03_representative_datasets.md}{\texttt{03\_representative\_datasets.md}} in \zenododataset{}.
For the final simulated data sets reads based on mixtures of the \plant{At} (TAIR10) core and chloroplast genome were generated with different ratios ( \num{0}:\num{1}, \num{1}:\num{10}, \num{1}:\num{100}, and \num{1}:\num{1000}).
Additionally, we generated data with different read lengths (\SIlist{150;250}{\basepair}). All data simulated contain exactly 2 million read pairs.

@@

\subsection*{Evaluation Criteria}
\subsubsection*{Computational Resources}
%Due to computational resources are limited, we evaluated the effect of different input parameters.
We recorded the mean and the peak CPU usage, the peak memory consumption, and the size of the assembly folder for each program. 
As input data, we used different data sets comprising \numlist{25000;250000;2500000} read pairs sampled from  our simulated reads.
%Therefore, the simulated perfect read files have been read repeatedly to achieve the required number of read pairs.
%\ak{etwas umständlich formuliert, hoffe ich hab es jetzt richtig formuliert und verstanden}
We used our docker image setup (\cref{tab:dockerimages}) \ak{check Nummerierung tables, oben war auch schon table 2} to run all assembly programs three times for each parameter setting.
The different settings combined different input data and different number of threads to use (\numlist{1;2;4;8}).
%\ak{in 'Our Setup' before, schreiben wir 10GiB und 4 threads, müsste dann vielleicht geändert werden}
Some programs will use more CPU threads than specified, therefore, the number of CPUs available have been fixed using the CPU option while running the docker run command.
For each assembly setting, we recorded the peak memory consumption, the CPU usage (mean and peak CPU usage) and the size of the folder where the assembly was calculated.
The values of CPU and memory usage have been obtained by docker.
The disk usage was estimated using the GNU tool \texttt{du}.
We used \texttt{GNU parallel} for queuing of the different settings \cite{Tange2011a}.

@@

\begin{itemize}
    \item Is the tool easy to install? 
    \item Is there a way to test the installation or a tutorial on how to use the tool? 
    \item Is there a good documentation on the parameter settings? 
    \item Is the tool maintained (issues answered, implementation of new features)? 
    \item Is the tool Open Source?
\end{itemize}
\newline
These questions were answered with \good{}, \ok{} or \bad{}, depending on the quality of the result. 
For example, a \good{} installation utilizes an automated package or dependency management like \texttt{apt}, CRAN, docker, etc.
An \ok{} installation procedure provides a custom script to install everything or at least lists all dependencies.
A \bad{} installation procedure fails to list important dependencies or produces errors, that prevent a successful installation without exhaustive debugging.

@@

\subsubsection*{Reproducibility}
\todo{lieber Consistency?}To ensure reproducibility of the obtained results, we randomly chose  \num{100} data sets, that in the previous runs resulted in outputs for most of the assemblers and run them again with the same parameters as before.
The resulting assemblies were scored again as described and the scores of the first and the second run were compared to each other
This information is important to assess the robustness of the different programs.

\subsubsection*{Time requirements}
In terms of run time, massive differences between the different tools have been observed. Apart from tool-specific differences, input data and number of threads had huge impact. The observed run times varied from a few minutes to several hours (\cref{fig:performance_runtime}).
%\ak{did I miss Figure 1 ?, need to check figure numbering in the end}
Some assemblies failed to finish within the time-limit we set (\SI{48}{\hour}). 
On average the longest time to generate the assemblies was taken by \ioga{} and \fp{} followed by \oa{} and \go{} the most time efficient tool was \ce{}, which on average is a little faster than \np{} and \cassp{}.
%% multithreading

Not all the tools were able to benefit from having access to multiple threads. Both \np{} and \oa{} take about the same time independent of being able to utilize \numlist[list-final-separator={, or }]{1;2;4;8} threads. \cassp{}, \ce{}, \go  and \fp{} all profit from multi-threading (\cref{fig:performance_runtime}, \cref{fig:performance_memory_cpu}).

\subsection*{Qualitative}
The user experience of most tools was evaluated as mainly \good{} (see table \ref{tab:resultsQual}). However, a few critique points remained.
Two minor dependencies were missing in the \go{} installation instructions and there was no test data available.
Additionally, an issue occurred when running it on a \plant{At} data set.
We are currently in the process of resolving this with the authors.

\subsection*{Quantitative}
\subsubsection*{Simulated data}
The only assembler obtaining perfect results according to our score for the simulated data sets is \go{} (\cref{fig:simulated}).
\ioga{} and \cassp{} showed the worst performance, being unable to fully assemble a single chloroplast out of \num{14} runs.
\np{} performed second best with scores above \num{80} for all data sets, only failing to resolve the contigs into one single circular chloroplast assembly.
The overall performance \ak{of all assemblers ?} is best, when the input data consists purely of chloroplast reads.
Only \ioga{} and \cassp{} failed to deliver any results under this scenario once.
In general, no clear correlation between either length of the input reads or the ratio of core vs chloroplast reads and the performance of the different assemblers can be observed.

>>>>

\input{biocon-stuff.tex}
\input{docker-stuff.tex}

@@

% List of docker images
\endlocaldefs

@@

\author[
    addressref={aff1,aff3},
    noteref={n1},
    email={jan.freudenthal@uni-wuerzburg.de}
    ]{\inits{JAF}\fnm{Jan A} \snm{Freudenthal}}
\author[
    addressref={aff1,aff2},
    noteref={n1},
    email={simon-pfaff@gmx.de}
    ]{\inits{SP}\fnm{Simon} \snm{Pfaff}}
\author[
   addressref={aff1,aff3},
   email={niklas.terhoeven@uni-wuerzubrg.de}
]{\inits{NT}\fnm{Niklas} \snm{Terhoeven}}
\author[
   addressref={aff1},
   email={arthur.korte@uni-wuerzubrg.de}
]{\inits{AK}\fnm{Arthur} \snm{Korte}}
\author[
   addressref={aff1,aff3,aff2 },
   noteref={n2},
   email={markus.ankenbrand@uni-wuerzubrg.de}
]{\inits{MJA}\fnm{Markus J} \snm{Ankenbrand}}
\author[
   addressref={aff1,aff2,aff4},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   noteref={n2},                        % id's of article notes, if any
   email={frank.foerster@uni-wuerzburg.de}   % email address
]{\inits{FF}\fnm{Frank} \snm{Förster}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstractbox}
\begin{abstract} % abstract
%One or two sentences providing a basic introduction to the field, comprehensible to a scientist in any discipline.
Chloroplasts are photosynthetic organelles in plant cells and contain their own genomic information.
That genome can be utilized in different scientific fields like phylogenetics or biotechnology.
%Two to three sentences of more detailed background, comprehensible to scientists in related disciplines.
Thus, different assemblers have been developed specialized in chloroplast assemblies.
Those assemblers often use the output of whole genome sequencing experiments as input.
Such sequencing data usually contain the complete chloroplast genome information, even if the sequencing aims for the core genome.
%One sentence clearly stating the general problem being addressed by this particular study.
Different assembly tools have never been systematically compared. %to each other concerning success rate or computational requirements.
%One sentence summarising the main result (with the words “here we show” or their equivalent).
Here we present a benchmark of seven chloroplast assembly tools, capable to succeed in more than \SI{60}{\percent} of real data sets.
%Two or three sentences explaining what the main result reveals in direct comparison to what was thought to be the case previously, or how the main result adds to previous knowledge.
Our results show significant differences between the tested assemblers in terms of generating whole chloroplast genome sequences and computational requirements.
Moreover, we suggest further development to improve user experience and success rate.
%One or two sentences to put the results into a more general context. 
In terms of reproducibility, we created docker images for each tested tool, which are available for the scientific community.
Following the presented guidelines, users are able to analyze and screen data sets for chloroplast genomes using only standard computer infrastructure.
%Two or three sentences to provide a broader perspective, readily comprehensible to a scientist in any discipline, may be included in the first paragraph if the editor considers that the accessibility of the paper is significantly enhanced by their inclusion. 
Thus large scale screening for chloroplasts as hidden treasures within genomic sequencing data is feasible.

@@

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Introduction}
\subsection*{General introduction and motivation}
Chloroplasts are essential organelles present in plant cells and the cells of some protists.
Chloroplasts enable the conversion of light energy into chemical energy via photosynthesis.
They harbor their own ribosomes and a circular DNA genome usually with a size between \SIrange{120}{160}{\kilo\basepair} \cite{palmer_1985}.
Because of this small size, the chloroplast genome has been an early target for sequencing.
The first chloroplast genome sequences were obtained as early as 1986 \cite{ohyama_chloroplast_1986,shinozaki_complete_1986}.
These early efforts elucidated the general genome organization and structure of the chloroplast DNA. 
Chloroplast genome content and structure are reviewed for example in \cite{wicke_evolution_2011,green_chloroplast_2011}.
Chloroplast genomes are widely used for evolutionary analyses \cite{martin_plastid_2010,xiao-ming_inferring_2017}, barcoding \cite{kress_use_2005,hollingsworth_dna_2009,de_vere_dna_2015}, and meta-barcoding \cite{bell_review_2016,deiner_environmental_2017}.
Interesting aspects of chloroplast genomes are their small size (\SIrange{120}{160}{\kilo\basepair},\cite{palmer_1985}), caused through endosymbiotic gene transfer \cite{martin_evolutionary_2002,timmis_endosymbiotic_2004} and the low number of \numrange{100}{120} genes that are still encoded on the chloroplast genome \cite{wicke_evolution_2011}.
Despite the overall high conservation of the genome sequence, there are striking differences in the gene content between different groups (e.g. the loss of the whole \genename{ndh} gene family in Droseraceae \cite{nevill_plastome-wide_2019}). Even more extreme evolutionary cases, where chloroplasts show a very low GC content and a modified genetic code are described \cite{su_novel_2019}.

@@

Structurally, two inverted repeats (IR\textsubscript{A} and IR\textsubscript{B}) of \SIrange{10}{76}{\kilo\basepair} divide the chloroplast genome into a large (LSC) and a small single copy (SSC) region \cite{palmer_1985}.
Those large inverted repeats complicate automated resolution with short read technologies\cite{Wang2018}.
Moreover, the existence of different chloroplasts within a single individual, and thus multiple different chloroplast genomes, have been described for different plants \cite{corriveau_1988,Chat2002,Scar2016}.
Although the origin and evolutionary importance of this phenomena ---called heteroplasmy--- are only poorly understood, the assembly of whole chloroplast genomes might be hindered.

Databases exist containing short read data for species where no reference chloroplast sequence is publicly available, eg.\ the Sequence Read Archive at NCBI \cite{sra2010}.
The availability of whole chloroplast genomes would enable large scale comparative studies \cite{tonti-filippini_what_2017}.
Additionally, reconstructed full chloroplast genomes have been used as super-barcodes \cite{coissac_barcodes_2016}, for biotechnology applications and genetic engineering \cite{daniell_chloroplast_2016}.

@@

Different strategies have been developed to assemble chloroplast genomes \cite{twyford_strategies_2017}.
In general, obtaining a chloroplast genome from WGS data requires two steps. First, the chloroplast reads have to be extracted from the mixed sequencing data. 
The second step is the assembly and resolution of the special circular structure including the inverted repeats.
The extraction of the reads can be achieved by mapping the reads to a reference chloroplast. \cite{Vinga2012}.
A different approach that does not perform alignments, relies on the higher coverage of chloroplast data in the whole genome sequencing data set\cite{Chan2013}.
Here, a $k$-mer analysis can be used to extract the most frequent reads.
An example for this is implemented in \ce{} \cite{j_ankenbrand_chloroextractor:_2018}. %\todo{ce ist hier der einzig explizit genannte Assembler, das wirkt wie bias :) entfernen? Nope... Steht klar ein Beispiel und wir haben nun mal den angesehen... Würde ich jetzt einmal so drin lassen. Alternativ nehmen wir noch einen dazu}
A third method combines both approaches by using a reference chloroplast as seed and simultaneously assembling the reads based on $k$-mers \cite{dierckxsens_novoplasty:_2017}.

%\ak{e.g. HomBlocks: A multiple-alignment construction pipeline for organelle phylogenomics based on locally collinear block searching, aber ist nicht inkluded, also können wir zitieren, müssen aber nicht}

@@

\subsubsection*{Time requirements}
In terms of run time, massive differences between the different tools have been observed. Apart from tool-specific differences, input data and number of threads had huge impact. The observed run times varied from a few minutes to several hours (\cref{fig:performance_runtime}).

Some assemblies failed to finish within the time-limit we set (\SI{48}{\hour}). 
On average the longest time to generate the assemblies was taken by \ioga{} and \fp{} followed by \oa{} and \go{} the most time efficient tool was \ce{}, which on average is a little faster than \np{} and \cassp{}.
%% multithreading

Not all the tools were able to benefit from having access to multiple threads. Both \np{} and \oa{} take about the same time independent of being able to utilize \numlist[list-final-separator={, or }]{1;2;4;8} threads. \cassp{}, \ce{}, \go  and \fp{} all profit from multi-threading (\cref{fig:performance_runtime,fig:performance_memory_cpu}).

\subsection*{Qualitative}
The user experience of most tools was evaluated as mainly \good{} (\cref{tab:resultsQual}). However, a few critique points remained.
Two minor dependencies were missing in the \go{} installation instructions and there was no test data available.
Additionally, an issue occurred when running it on a \plant{At} data set.
We are currently in the process of resolving this with the authors.

\subsection*{Quantitative}
\subsubsection*{Simulated data}
The only assembler obtaining perfect results according to our score for the simulated data sets is \go{} (\cref{fig:simulated,tab:scores_simulated}).
\ioga{} and \cassp{} showed the worst performance, being unable to fully assemble a single chloroplast out of \num{14} runs.
\np{} performed second best with scores above \num{80} for all data sets, only failing to resolve the contigs into one single circular chloroplast assembly.
The overall performance is best, when the input data consists purely of chloroplast reads.
Only \ioga{} and \cassp{} failed to deliver any results under this scenario once.
In general, no clear correlation between either length of the input reads or the ratio of core vs chloroplast reads and the performance of the different assemblers can be observed.

\subsubsection*{Real data sets}
Concerning the performance of the assemblers on the real data sets, we were able to observe considerable differences in the median score (\cref{fig:swarmplot}).
The highest scores were achieved by \go{} with a median of \num{99.7} and \num{199} circular assemblies out of a total of 356 assemblies that resulted in an output (\cref{tab:scores_real}).
The performance of \go{} is followed by \fp{}, \np{}, \oa{}, and \ce{}. \fp{} is outperforming the latter two slightly in terms of score, with twice as much \num{114} perfectly assembled chloroplast genomes (\np{} produced \num{66} and \oa{} \num{55} circular genomes).
\ioga{} and \cassp{} were both not able to assemble a circular, single-contig genome (\cref{tab:scores}), consequently resulting in the lowest mean and median scores (\cref{fig:upset}).

\subsubsection*{Consistency}
Consistency was tested by re-running assemblies and comparison of the scores of two assemblies (\cref{fig:consistency}).
Replicates that did not produce an output were manually scored as \num{0}. 
\go{} was the only tool that succeeded in obtaining similar scores for all assemblies, without producing and completely unsuccessful assemblies for this subset of data.
Except for \fp{} all the other tools had at least one assembly that was unsuccessful in one run, but produced an output in the other.
%\ak{oben steht \go{} war als einziges immer gut, dann kommt except for \fp{} ?? does not seem to be consistent}
Notably \ioga{} appears to have a tendency to perform differently in independent runs.
Here, more than \SI{10}{\percent} of the assemblies failed in one run only.

Both \fp{} and \np{} tend to have minor changes in the assembly when the overall performance is comparably well, leading to the arrow-shaped scatter plots.
%\ak{link to figure scatter plots is missing}
\ce{} and \cassp{} appear to be the most robust assemblers, having only few deviations between the two runs.

\subsection*{Guidelines for the end-user}
Given these results, our recommendation is to use \go{} as default option, and in case of failure \fp{} as backup solution.
%\ak{what if both lead to qualitative different results?, always trust \go{} , or could one compute a high-confidence assembly by combining different results ?, might be a point for future directions.}
If both programs fail, it is sensible to re-run \fp{} and additionally try \np{} and \oa{}.
This procedure maximizes the chance to effectively and efficiently recover the circular chloroplast genome from mixed genomic data.
If none of these four assemblers produce sensible results, a reference guided approach and tweaking of the default parameters, might be the solution.
Here, it is not possible to provide general guidelines, as the procedure will differ for different data sets.
For an automated approach, running \go{} and \fp{} in parallel appears to be a good trade-off between success rate and use of resources.

\subsection*{Ideas for future development}
For further experiments, combining different components from different tools might be a promising approach.
For example, read scaling from \ce{} followed by an assembly by \go{} and finally the structural resolution with \fp{} could be a promising approach, combing the respective strength of the different tools.

Another important feature of software is a comprehensive documentation, which needs to be up-to-date and maintained.
Additionally, software authors could improve the usability based on suggestions from their users.

Finally, all tools should improve their integrated guessing of default parameters, as many users avoid fine tuning of those, especially, for larger screening approaches.
Last, as sequencing technology is developing fast (eg.\ PacBio or nanopore), tools need to be updated to not become obsolete.
But the hope would be that with ongoing software development and improved sequencing technologies, the generation of whole chloroplast assemblies from any species will become a routine technique.

\section*{Conclusion}
The main assumption for our study to benchmark different chloroplast assembly tools, is that whole genome sequencing data are also a promising source for chlorplast assemblies.
Our benchmark shows that ~\SI{60}{\percent} of the data sets without available chloroplast genome, have been assembled by at least one of the tools we analysed.
Still, even with simulated (aka``perfect'') data, not all tools succeeded in generating complete chloroplast assemblies.
Therefore, we determined the strengths and weaknesses of the specific tools and provided guidelines for the users.
However, it might be necessary, to combine different methods or manually explore the parameter space, to obtain reliable results if a single run seems not sufficient.
Ultimately, large scale studies reconstructing hundreds or thousands of chloroplast genomes are now feasible using the currently available tools. 
%Outlook: next step in the pipeline is annotation. There is a similar problem: many tools, which one to use? 
%A large scale study reconstructing chloroplasts throughout SRA (or \num{1001} genome project) is feasible and promising using the existing tools (estimated runtime, number of novel chloroplasts).
%A meta-tool that automatically combines the different programs could be developed.
%The benchmarking study could be augmented to automatically select the closest reference from the database and provide that to any tool that can take advantage of references.

\section*{Methods}
\subsection*{Data availability}
Source code for all methods used is available at \cite{github-benchmark-repo} and archived in zenodo under \zenodorepo{}.
All docker images are published on \cite{dockerhub-benchmark} and are named with a leading \texttt{benchmark\_} (\cref{tab:dockerimages}).

@@

\subsection*{Our Setup}
We want to use a minimum of different parameter settings for all assembly programs to enable a fair comparison. 
Therefore, we decided to specify that all programs have to work based on two input files, representing a data set's forward (\texttt{forward.fq}) and reverse  (\texttt{reverse.fq}) sequence file in FASTQ format.
Depending on the assembler, output files with different names and locations are generated.
Those different files are copied and renamed to ensure that each assembly approach produces the same output file (\texttt{output.fa}). 
Additionally, we set an environment variable for all programs to control the number of allowed threads.
All three requirements (defined input file names, defined output file name, thread number control via environment variable) are ensured by a simple wrapper script (\texttt{wrapper.sh}).
Finally, for a maximum of reproducibility all programs have been bundled into individual docker images based on a central base image which provides all the required software. Those docker images were used for the recording of the consumption of computational resources.
Furthermore, all our docker images have been converted into singularity containers for the quantitative measurement on simulated and real data sets.
Singularity container were built from docker images for usage on a HPC-environment using Singularity v.2.5.2 \cite{kurtzer2017singularity}
All singularity containers were run on %the Julia-HPC 
Intel® Xeon® Gold \num{6140} Processors using a Slurm workload manager version 17.11.8 \cite{Jette02slurm}. Assemblies were run on \num{4} threads using \SI{10}{\gibi\byte} RAM with a time limit of \SI{48}{\hour}. 
\subsection*{Data}
\subsubsection*{Simulated}
To avoid suffering from sequencing errors and biological variances, we simulated perfect reads based on the \plant{At} (TAIR10) chloroplast assembly \cite{tair10}.
We used a sliding window approach with \texttt{seqkit} \cite{seqkit}. The exact commands are documented in \href{https://github.com/chloroExtractorTeam/benchmark/blob/master/03_representative_datasets.md}{\texttt{03\_representative\_datasets.md}} in \zenododataset{}.
For the final simulated data sets reads based on mixtures of the \plant{At} (TAIR10) core and chloroplast genome were generated with different ratios ( \num{0}:\num{1}, \num{1}:\num{10}, \num{1}:\num{100}, and \num{1}:\num{1000}).
Additionally, we generated data with different read lengths (\SIlist{150;250}{\basepair}). All data simulated contain exactly 2 million read pairs.

@@

\subsection*{Evaluation Criteria}
\subsubsection*{Computational Resources}
%Due to computational resources are limited, we evaluated the effect of different input parameters.
We recorded the mean and the peak CPU usage, the peak memory consumption, and the size of the assembly folder for each program. 
As input data, we used different data sets comprising \numlist{25000;250000;2500000} read pairs sampled from  our simulated reads.
%Therefore, the simulated perfect read files have been read repeatedly to achieve the required number of read pairs.
We used our docker image setup (\cref{tab:dockerimages}) to run all assembly programs three times for each parameter setting.
The different settings combined different input data and different number of threads to use (\numlist{1;2;4;8}).

Some programs will use more CPU threads than specified, therefore, the number of CPUs available have been fixed using the CPU option while running the docker run command.
For each assembly setting, we recorded the peak memory consumption, the CPU usage (mean and peak CPU usage) and the size of the folder where the assembly was calculated.
The values of CPU and memory usage have been obtained by docker.
The disk usage was estimated using the GNU tool \texttt{du}.
We used \texttt{GNU parallel} for queuing of the different settings \cite{Tange2011a}.

@@

\begin{itemize}
    \item Is the tool easy to install? 
    \item Is there a way to test the installation or a tutorial on how to use the tool? 
    \item Is there a good documentation on the parameter settings? 
    \item Is the tool maintained (issues answered, implementation of new features)? 
    \item Is the tool Open Source?
\end{itemize}

These questions were answered with \good{}, \ok{} or \bad{}, depending on the quality of the result. 
For example, a \good{} installation utilizes an automated package or dependency management like \texttt{apt}, CRAN, docker, etc.
An \ok{} installation procedure provides a custom script to install everything or at least lists all dependencies.
A \bad{} installation procedure fails to list important dependencies or produces errors, that prevent a successful installation without exhaustive debugging.

@@

\subsubsection*{Consistency}
To ensure consistency of the obtained results, we randomly chose  \num{100} data sets, that in the previous runs resulted in outputs for most of the assemblers and run them again with the same parameters as before.
The resulting assemblies were scored again as described and the scores of the first and the second run were compared to each other
This information is important to assess the robustness of the different programs.
